Paradigms in Physics 2.0

### 2017-06-29 paper.tex
first submitted draft

\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}

@@

\section{Paradigms 2.0 process}
We began the Paradigms~2.0 process in the Winter of 2016, having in
the previous year obtained faculty agreement, and support from our
Department head.  The process was spearheaded by a committee of four
(the authors DR, EG, EM, and CAM, with EvZ present at meetings
documenting our process).  During the Winter quarter, we informed the
community of our process through a colloquium, and collected student
and faculty perspectives on the existing curriculum.  We then
interviewed faculty who recently taught each of our existing courses
to document topics that were currently covered.  This resulted in a
total of $\sim$700 index cards in $\sim$30 stacks, with each stack
corresponding to one course, and each card describing a topic,
color-coded as in Fig.~\ref{fig:schedule}.

@@

While contemporary challenges determine the narrative flow of the
course, physics concepts and physical reasoning skills are the main
substance. Students are introduced to thermodynamics, statistical
mechanics, electromagnetic radiation, quantum mechanics, and modern
experimental physics---in each case motivated by one or more
``contemporary challenges.'' Together with new physics concepts,
students are given new reasoning tools, such as the equipartition
theorem, the quantum-classical correspondence principle, order of
magnitude calculations, simplifying assumptions, and numerical
integration. Mathematically rigorous derivations are only briefly
discussed in class, with additional details offered in online video
format. In summary, we emphasize the physics concepts and reasoning
skills that allow professional physicists to quickly/quantitatively
make an initial assessment of a complex problem.

The students engage with the material by solving daily quiz questions
in class, solving weekly homework problems, performing a modern
physics experiment, and writing a term paper. The daily quiz questions
help clarify the goal of each lecture. The weekly homework questions
are directly related to contemporary challenges requiring a variety
of estimation techniques. The modern physics experiment, performed
with a low cost kit (LEDs, diffraction grating, rulers, volt meter
etc.) emphasizes a quantitative approach to observation and explores
issues of experimental uncertainty. Lastly, the term paper allows
students to dive deeper into a particular contemporary challenge. The
term paper tests students' ability to make their own initial
assessment of a complex problem, and communicate their ideas with
diagrams, graphs and estimations.

@@

\section{Summary}
We have developed a significant change to the Paradigms
curriculum.  This change focuses on the first two big ideas underlying
the Paradigms project: (1) close attention to content ordering, and (2)
developing a consensus and understanding of the curriculum among our
faculty members.  Our commitment to using active engagement in our
upper-division curriculum remains unchanged.  While the specific
content ordering we developed may be interesting, the \emph{process} by
which we reached it and developed a faculty consensus is at the
essence of what makes the Paradigms in Physics program distinctive.
Emily van Zee, a researcher with experience in ..., 
\corinnesays{Emily, please add a short description of yourself} 
has completed a retrospective study of the process we used to develop the 
original Paradigms program and we have used her results extensively in 
designing and implementing the Paradigms~2.0 process.  She has continued 
to observe, study, and document our work as we have begun implementing 
the revised curriculum~\cite{lessonslearned}.
The process of developing faculty understanding of the Paradigms is
ongoing, as we are having professors new to the department shadow
experienced faculty as they teach \emph{Paradigm} course, prior to teaching
the same course themselves.

>>>>

\newcommand\corinnesays[1]{\textcolor{red}{[\sc CAM: {\em#1}]}}
\newcommand\davidsays[1]{\textcolor{blue}{[\sc DR: {\em#1}]}}
\renewcommand\davidsays[1]{}

@@

\section{Paradigms 2.0 process}
Like the process that led to Paradigms 1.0, the Paradigms 2.0 process is a ``shared vision'' type of change \cite{henderson2010}. Changes to the course structures of the physics curriculum emerged from discussions among the members of the department. We began the Paradigms~2.0 process in the Winter of 2016, having in
the previous year obtained faculty agreement, and support from our
Department head.  The process was spearheaded by a committee of four
(the authors DR, EG, EM, and CAM, with EvZ present at meetings
documenting our process).  During the Winter quarter, we informed the
community of our process through a colloquium, and collected student
and faculty perspectives on the existing curriculum.  We then
interviewed faculty who recently taught each of our existing courses
to document topics that were currently covered.  This resulted in a
total of $\sim$700 index cards in $\sim$30 stacks, with each stack
corresponding to one course, and each card describing a topic,
color-coded as in Fig.~\ref{fig:schedule}.

@@

While contemporary challenges determine the narrative flow of the
course, physics concepts and physical reasoning skills are the main
substance. Students are introduced to thermodynamics, statistical
mechanics, electromagnetic radiation, quantum mechanics, and modern
experimental physics---in each case motivated by one or more
``contemporary challenges.'' Together with new physics concepts,
students are given new reasoning tools, such as the equipartition
theorem, the quantum-classical correspondence principle, order of
magnitude calculations, simplifying assumptions, and numerical
integration. Mathematically rigorous derivations are only briefly
mentioned in class. Instead we emphasize the physics concepts and reasoning
skills that allow professional physicists to quickly/quantitatively
make an initial assessment of a complex problem.

% The students engage with the material by solving daily quiz questions
% in class, solving weekly homework problems, performing a modern
% physics experiment, and writing a term paper. The daily quiz questions
% help clarify the goal of each lecture. The weekly homework questions
% are directly related to contemporary challenges requiring a variety
% of estimation techniques. The modern physics experiment, performed
% with a low cost kit (LEDs, diffraction grating, rulers, volt meter
% etc.) emphasizes a quantitative approach to observation and explores
% issues of experimental uncertainty. Lastly, the term paper allows
% students to dive deeper into a particular contemporary challenge. The
% term paper tests students' ability to make their own initial
% assessment of a complex problem, and communicate their ideas with
% diagrams, graphs and estimations.

@@

\section{Summary}
We have developed a significant change to the Paradigms
curriculum.  This change focuses on the first two big ideas underlying
the Paradigms project: (1) close attention to content ordering, and (2)
developing a consensus and understanding of the curriculum among our
faculty members.  Our commitment to using active engagement in our
upper-division curriculum remains unchanged.  While the specific
content ordering we developed may be interesting, the \emph{process} by
which we reached it and developed a faculty consensus is at the
essence of what makes the Paradigms in Physics program distinctive.
One of the authors (EvZ) have completed a retrospective study of the 
process we used to develop the 
original Paradigms program and we have used her results extensively in 
designing and implementing the Paradigms~2.0 process.  She has continued 
to observe, study, and document our work as we have begun implementing 
the revised curriculum~\cite{lessonslearned}.
The process of developing faculty understanding of the Paradigms is
ongoing, as we are having professors new to the department shadow
experienced faculty as they teach \emph{Paradigm} courses, prior to teaching
the same course themselves.
